---
title: "Influenza Bern"
author: "Katarina Matthes"
date: "`r Sys.Date()`"
output:
  rmdformats::robobook:
    code_folding: show
    self_contained: true
    highlight: pygments
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Methods
* Bayesian approach
* INLA (Bayesian inference for Latent Gaussian Models)
* Influenza cases for $475$ municipalities of canton of Bern for years 1918 - 1926
* Data $Y$ is given as the total number of influenza in fixed municipalities and in each year
* Standard Poisson likelihood to model the counts $$ y_i \mid \eta_i\sim Po\left(E_i \exp(\eta_i)\right),$$
where $E_i$ is the ``population at risk'' in municipality $i$.
<br />
<br />
* classical disease mapping model; BYM model (Besag, York and Mollie proposed it)
* The log relative risk, $\bf{\eta} = (\eta_1, \dots, \eta_n)^T$, is thus decomposed
  into $$\bf{\eta} = \mu + \bf{u} + \bf{v} $$
  +  $\mu$ is the overall intercept
  +  $\bf{u}$ is a random effect with spatial structure following the Besag model
  +  $\bf{v}$ represents a non-spatial overdispersion
* ${\bf u}$ is "besag" modelled spatially structured with smoothing parameter $\kappa_u$.
* ${\bf v}$ is unstructured with precision parameter $\kappa_v$, i.e.
$\bf{v} \sim \mathcal{N}(0, \kappa_v^{-1}I)$.
* The precision terms $\kappa_v$ and $\kappa_u$ are assigned the default
gamma prior distributions of INLA
$$
\begin{aligned}
  \kappa_u & \sim \textrm{Gamma}( \alpha_u, \beta_u ), \\
  \kappa_v & \sim \textrm{Gamma}( \alpha_v, \beta_v ).
\end{aligned}
$$
* The default values are $\alpha_u = \alpha_v = 1$ and $\beta_u = \beta_v = 0.00005$.
<br />
<br />

```{r, echo=TRUE,message=FALSE, eval=FALSE}
formula <- Num ~ 1 + offset(log(Population)) +
  f(Region, model="bym", graph="Gemeinde_Inla", scale.model = TRUE)

 inla.mod <- inla(formula,
                   data=dat.bern,
                   family = "zeroinflatednbinomial0",
                   control.family = control.family,
                   control.compute = list(config = TRUE),
                   control.mode = list(restart = TRUE),
                   control.predictor = list(compute = TRUE, link = 1))
```

* 1000 samples from the posterior distribution

# Spanish flu

## Maps
Fitted values

```{r, echo=FALSE,message=FALSE,fig.height = 8,  fig.width = 10}
function_maps_wave(Wave_Inf=1)
function_maps_wave(Wave_Inf=2)
function_maps_wave(Wave_Inf=3)
```

## Hotspot
Fitted values

```{r, echo=FALSE,message=FALSE,fig.height = 8,  fig.width = 10}
function_hotspot(Wave_Inf=1)
function_hotspot(Wave_Inf=2)
function_hotspot(Wave_Inf=3)
```

